---
title: "Hotel Reservation Prediction"
output: pdf_document
---

# Exploratory Data Analysis

### Load dataset

```{r}
df <- read.csv("hotel_reservation.csv")
```

### Dimensionality

```{r}
dim(df)
```

### Structure

```{r}
str(df)
```

### Summary

```{r}
summary(df)
```

### Import library

```{r}
library(dplyr)
library(ggplot2)
```

### Calculate the number of guests from different countries

```{r}
country_guests <- df%>%
  # Group the dataset by country
  group_by(country)%>%
  # Replace null as OTHERS for better understanding
  mutate(country=ifelse(is.na(country)|country=="","OTHERS",country))%>%
  # Calculate the sum of guests
  summarise(guests=sum(adults+children+babies),
            canceled=sum((adults+children+babies)*(is_canceled=="1")),
            not_canceled=sum((adults+children+babies)*(is_canceled=="0")))%>%
  # Show the result of countries in the order from the most to least guests
  arrange(desc(guests))
print(country_guests)
```

### Visualize the number of guests from different countries

```{r fig.width = 10, fig.height=15}
ggplot(data=country_guests,mapping=aes(x=guests,y=reorder(country,guests),fill=reorder(country,guests)))+
  # Adjust the countries in order of guests
  geom_col()+
  # Set title
  labs(y="country",title="Number of guests from different countries")+
  theme(plot.title=element_text(size=24,hjust=0.5),
        # Don't show legend
        legend.position="none",
        # Set background
        panel.grid.major=element_line(color="grey60",linewidth=0.25),
        # Change text size
        axis.text.y=element_text(size=10),
        axis.text.x=element_text(size=14),
        axis.title.y=element_text(size=20),
        axis.title.x=element_text(size=20))
```

#### Analysis: 

There are a total of 119 countries in the picture, and when drawing the picture, we found that some of them have empty values, so they are named after "OTHERS". From the graph, it can be observed that the top 10 countries with the highest number of residents are Portugal, the United Kingdom, France, Spain, Germany, Italy, Ireland, Brazil, Belgium, and the Netherlands. Among them, Portugal has the highest number of hotel guests, surpassing the second ranked country by about three times. And all the top 10 countries are from Europe. On the contrary, we can also observe from the visualization that only 1-2 people from 21 countries are staying in these two hotels, most of which come from West Africa and Central America.

```{r}
library(maps)
library(countrycode)

# Obtain the country border information
world_map <- map_data("world")
country_guests$country_name <- countrycode(country_guests$country, "iso3c", "country.name")

# Assigns "China"/"Timor-Leste" to the country_name attribute of a record 
# whose country attribute is "CN"/"TMP" in the dataset
country_guests <- country_guests %>%
  mutate(country_name = ifelse(country == "CN", "China", country_name),
         country_name = ifelse(country == "TMP", "Timor-Leste", country_name))

# Merge data
merged_data <- merge(world_map, country_guests, 
                     by.x="region", by.y="country_name", 
                     all.x = TRUE)

# Create Heat Map
library(ggplot2)

ggplot(merged_data, 
       aes(x=long, y=lat, 
           group=group, 
           fill=guests)) +
  geom_polygon() +
  scale_fill_gradient(low="pink", high="red", name="Total Guests") +
  theme_void() +
  labs(title="Guests Heatmap by Country")+
  theme(plot.title=element_text(hjust=0.5),
        legend.text=element_text(size=8),
        legend.title=element_text(size=10))
```

#### Analysis:

In order to clearly show the geography distribution of the countries, we also draw the map, which showed that most of the customers are from Europe. Also, some of the countries that have high volume of customers are located in Africa.

### Plot the top 10 countries with the largest number of guests

```{r}
library(tidyr)
# Get data of top 10 countries
top10_country <- head(country_guests,10)%>%
  select(country,not_canceled,canceled)%>%
  gather("type","guests",-country)
# Plot
ggplot(data=top10_country,mapping=aes(x=reorder(country,-guests),y=guests,fill=type))+
  # Set color of bars
  geom_bar(stat="identity")+
  # Set title
  labs(x="country",title="Number of guests (Top10 countries)")+
  # Reset legend for better understanding
  scale_fill_manual(breaks=c("not_canceled","canceled"),
                    labels=c("not canceled","canceled"),
                    # Set color for bars
                    values=c("lightblue","pink"))+
  theme(plot.title=element_text(hjust=0.5),
        # Set background
        panel.grid.major=element_line(color="grey60",linewidth=0.25))
```

#### Analysis: 

As shown in the above figure, there are significant differences in the number of customers from different countries, making it difficult to display details in one image. Therefore, we selected the top 10 countries with the highest number of customers to observe customer distribution and cancellation numbers. Although Portugal has the largest number of customers, it also has a high cancellation rate, with a cancellation rate greater than 50%. Among the countries with the second to tenth highest number of customers, although there are also some cancellations, the proportion of cancellations is relatively small compared to Portugal.

### Create an order for month, for latter visualization

```{r}
month_in_order <- c("January", "February", "March", "April", "May", "June",
                    "July", "August", "September", "October", "November", "December")
```

### Calculate the average number of nights the guests stayed per month

```{r}
stay_nights_month <- df%>%
  # Group data by two attributes
  group_by(arrival_date_month,is_canceled)%>%
  # Calculate the result
  summarise(nights=sum(stays_in_weekend_nights+stays_in_week_nights)/n(),
            .groups="drop_last")%>%
  # Change the order of arrival_date_month
  mutate(arrival_date_month=factor(arrival_date_month,level=month_in_order))%>%
  # Let the result show in the order of month
  arrange(is_canceled,arrival_date_month)

# Print dataset
print(stay_nights_month)
```

### Visualize the average number of nights the guests stayed per month

``` {r}
# Plot
ggplot(data=stay_nights_month,mapping=aes(x=arrival_date_month,y=nights,
                                          fill=as.factor(is_canceled)))+
  geom_bar(stat="identity",position="dodge")+
  # Set x label and title
  labs(x="month",title="Average number of nights the guests stayed per month")+
  # Reset legend for better understanding
  scale_fill_manual(name="type",
                      breaks=c("0","1"),
                      labels=c("not_canceled","canceled"),
                      # Set color for bars
                      values=c("lightblue","pink"))+
  # Change text size and angle
  theme(plot.title=element_text(size=12,hjust=0.5),
        axis.text.x=element_text(angle=40,hjust=1),
        # Set backgraound
        panel.grid.major=element_line(color="grey60",linewidth=0.25))
```

#### Analysis: 

We arranged the data months in order and calculated the average number of nights spent in hotels per month for customers who did not cancel their orders, and also for those who cancelled their orders.

From the graph, the average number of days a customer stays in a hotel per month fluctuates with each month, whether they have cancelled or not. Among them, the average number of nights for both canceled and not canceled from July to August is the highest, around 4 days. From the graph, it can be observed that the average length of stay in the hotel is about 3.5 days.

Customers who have not cancelled have a stable trend of first increasing and then decreasing in their stay days from January to August and August to December, while customers who have cancelled their stay have a higher average stay days in January, March, July, August, and December.


### Calculate the number of guests per month for both Resort Hotel and City Hotel

```{r}
guests_per_month <- df%>%
  # Group data by 3 attributes
  group_by(hotel,arrival_date_month,is_canceled)%>%
  # Calculate the number of guests
  summarise(guests=sum(adults+children+babies),.groups = "drop_last")%>%
  # Change the order of arrival_date_month
  mutate(arrival_date_month=factor(arrival_date_month,level=month_in_order))%>%
  # Arrange the result in an order
  arrange(is_canceled,arrival_date_month,hotel)
print(guests_per_month)
```

### Visualize the number of guests per month for both Resort Hotel and City Hotel

``` {r}
# Create labels for changing the sub titles
labels <- c("0"="Not canceled","1"="Canceled")
# Plot
ggplot(data=guests_per_month,mapping=aes(x=arrival_date_month,y=guests,fill=hotel))+
  geom_bar(stat="identity",position="dodge")+
  # Create sub plots
  facet_grid(~as.factor(is_canceled),labeller=as_labeller(labels))+
  # Set x label and title
  labs(x="month",title="Number of guests per month")+
  # Set color for bars
  scale_fill_manual(values=c("pink","lightblue"))+
  # Change text size and angle
  theme(plot.title=element_text(size=12,hjust=0.5),
        axis.text.x=element_text(angle=40,hjust=1,size=6),
        # Set background
        panel.grid.major=element_line(color="grey60",linewidth=0.25))
```


#### Analysis: 

In order to calculate the average number of guests per month, we divided customers into cancelled orders and non cancelled orders for observation. It can be observed that there is a similar trend in the number of customers who have not cancelled orders and who have cancelled orders every month, both showed an upward trend from January to August, while the number of customers checking in from August to December showed a downward trend.

By further dividing the images into City hotels and Resort hotels for observation, it can be observed that the average number of people staying at City hotels exceeds that of Resort hotels per month.

We speculate that the number of people staying at different hotels each month has the same trend, and there is a certain correlation with the month.

### Calculate the average hotel price (adr) of each month for both Resort Hotel and City Hotel

```{r}
# Get data of all uncancelled orders
adr_per_month <- df[which(df$is_canceled=="0"),]%>%
  # Group data by hotel name and arrival month
  group_by(hotel,arrival_date_month)%>%
  # Calculate adr per month for each hotel
  summarise(price=sum(adr*(stays_in_weekend_nights+stays_in_week_nights))
            /sum(stays_in_weekend_nights+stays_in_week_nights),.groups="drop_last")%>%
  # Change the order of arrival_date_month
  mutate(arrival_date_month=factor(arrival_date_month,level=month_in_order))%>%
  # Let the result show in the order of month and hotel
  arrange(arrival_date_month,hotel)
print(adr_per_month)
```

### Visualize the average hotel price (adr) of each month for both Resort Hotel and City Hotel

```{r}
# Plot
ggplot(data=adr_per_month,mapping=aes(x=arrival_date_month,y=price,color=hotel,group=hotel))+
  geom_line()+
  # Set x label and title
  labs(x="month",y="adr",title="Average hotel price (adr) of each month")+
  # Set color for bars
  #scale_color_manual(values=c("pink","lightblue"))+
  # Change text size and background
  theme(plot.title=element_text(size=12,hjust=0.5),
        axis.text.x=element_text(angle=30,hjust=1),
        panel.grid.major=element_line(color="grey60",linewidth=0.25))
```

#### Analysis: 

From the graph, it can be seen that the average monthly price of City hotel is relatively stable, with a steady upward trend from January to May. However, after June, although the price fluctuated slightly, the overall trend showed a downward trend. The average price reached its highest level in May, around 125 yuan per night.

The prices of the Resort Hotel showed a clear upward trend from January to August, especially with a significant increase in prices from July to August, and gradually decreasing after that.

Compared to City Hotel and Resort Hotel, it can be seen that except for July and August, the prices of Resort Hotel are significantly higher than those of City Hotel, while the prices of Resort are lower at other times.

From the graph, we speculate that there is a certain correlation between the average price of hotels and the month.

### The detailed and descriptive results of analysis are shown beneath each graph, here are some general analysis conclusion.

1. The majority of customers come from European countries, and you can focus on observing customer profiles of European customers, such as behavioral habits, stay days, etc. This helps the hotel identify potential customers who may cancel orders.

2. Hotels can analyze customer behavior to understand why the countries with the highest proportion of customers have higher room cancellations, in order to develop appropriate strategies to retain potential customers.

3. From July to August, customers have the highest average number of nights staying in hotels, with the highest average hotel prices per month. It can also be observed that the number of customers who cancel orders during this time period is also the highest. If some means can be used to retain customers, the hotel will gain greater profits.

4. We consider July to August as the peak season for tourism, as the hotel has the highest number of guests. Due to the lower monthly average price of City Hotel compared to Resort Hotel, the number of people staying at City Hotel is much higher than that at Resort Hotel. However, the cancellation rate of City Hotel is also higher than that of Resort Hotel, so City Hotel can analyze the specific reasons for high occupancy and cancellation rates.

# Data Pre-processing

### Check the missing values

```{r}
df[df == ""] <- NA
colSums(is.na(df))
```

### Handle the missing values. 

Firstly, it can be observed from the above statistics that there are only three attributes with missing values, namely country, agent, and company. Therefore, we have different processing methods for these different attributes.

Regarding country, we found that the proportion of null values in the dataset is very small, only 92/17908(0.51%). So, we have decided to directly delete these missing values. After removing the null value for country, there are still 17816 pieces of data in the dataset.

```{r}
# Remove countries with missing values
df <- subset(df, !is.na(country))
```

About agents and companies. In the dataset paper "Hotel Demand Dataset" [1], it was described that there are no NULL values for agent and company, as their NULL values represent a special category that does not come from agent and company reservations. So, we plan to fill in the empty values of these companies and agents into a new category of 0.

```{r}
# The missing values of agent and company are filled to 0
df$agent <- ifelse(is.na(df$agent), 0, df$agent) 
df$company <- ifelse(is.na(df$company), 0, df$company)
```


### Check duplicates and remove

Observing the dataset and according to the given variable description.txt file, we found that in the attribute 'meal', SC and Undefined represent the same category, both referring to 'no meal package'. Therefore, we decided to merge these two categories into one and check for redundant data.

```{r}
# Replace Undefined with SC
df$meal <- ifelse(df$meal == "Undefined", "SC", df$meal)
```

We found a total of 2822 redundant dayta in the dataset, with 14994 data remaining after deletion.

```{r}
# remove duplicates
df <- unique(df)
```


### Data distribution

```{r fig.width = 15, fig.height=3}
par(mfrow = c(1, 4))
variables <- colnames(df) 
for (variable in variables) {
  x <- df[[variable]]
  barplot(table(x),
          xlab = variable, col = "lightgreen",
          ylim = c(0, max(table(x)) + 100),
          main = paste("Data Distribution of", variable),
          cex.names = 0.5
  )
  box()
}
```

### Check and remove outliers

When dealing with outliers, we only observe and process numerical type attributes. After observation and consideration, we ultimately believe that the attributes that may have outliers are adults, children, babies, booking_changes and adr. The following will provide a detailed explanation of our judgment method and processing results.

Based on the data distribution graph drawn above, we can observe that the distribution of adults is mostly between 0-3 people. Based on the data distribution results and common sense, we believe that 0 adult hotel rooms with 5 or more people in one room are abnormal. So we decided to delete these situations. Based on the observation of the data distribution map of children and babies, we believe that although it is normal for the number of children and babies to be 0, it is abnormal for children or babies with more than 5 rooms to be occupied in one room. Also, we consider that for people who kept changing his/her reservation for more than 5 times are also abnormal according to the common sense. Therefore, such results are considered as outliers for deletion.

These commons sense for deletion also been checked by the data distribution plot, and those points that considered as ouliers are obviously away from most points.

The final amount of data to be deleted is 76, leaving 14918 rows of data remaining.

```{r}
# Remove outliers of adults, children, babies, booking changes
df <- df[!(df$adults == 0 | df$adults >= 5), ]
df <- df[!(df$children >= 5), ]
df <- df[!(df$babies >= 5), ]
df <- df[!(df$booking_changes > 5), ]
```

The following is an analysis of ADR outliers. Due to the fact that the price of the hotel is 0, which is not reasonable data based on common sense, we will delete it first.

```{r}
# Outliers of adr
# Remove inconsistent data
df <- df[df$adr != 0, ]
```

From T1-5, it can be seen that there is a certain price difference between City Hotel and Resort Hotel. In July August, the prices of Resort Hotel were significantly higher than those of City Hotel, while in other months they were the opposite. Therefore, we believe that the price difference between City Hotel and Resort Hotel should also be considered when selecting outliers. So when analyzing the outliers of ADR, we decided to have a group discussion.

When detecting outliers, we considered the z-score method and the boxplot method. Due to the fact that the z-score detection of outliers requires a rough normal distribution of the data, and it was observed through the data distribution map that ADR does not meet this condition, we ultimately chose the box plot method. Then we draw the box diagram of ADR for two hotels separately. For data outside the upper and lower limits of the box plot, delete between them.

Ultimately, the Resort Hotel is considered to have 131 outliers, while the City Hotel is considered to have 222 outliers. After removing the outliers, there are still 14347 pieces of data left.

```{r}
# Get Hotel type
hotel_types <- unique(df$hotel)

par(mfrow = c(1, 2))

# Box plot of different hotel types
for (hotel_type in hotel_types) {
  adr_df <- df$adr[df$hotel == hotel_type]
  boxplot(adr_df,
          ylab = "ADR",
          main = paste("ADR Boxplot for", hotel_type))
}

# Create data frame
outliers_df <- data.frame()

# Calculate outliers for adr
for (hotel_type in hotel_types) {
  hotel_df <- df[df$hotel == hotel_type, ]
  
  Q1 <- quantile(hotel_df$adr, 0.25)
  Q3 <- quantile(hotel_df$adr, 0.75)
  IQR <- Q3 - Q1
  # The threshold is set to 1.5
  Minimum <- Q1 - 1.5 * IQR
  Maximum <- Q3 + 1.5 * IQR
  
  # Judge outliers
  outliers <- which(hotel_df$adr < Minimum | hotel_df$adr > Maximum)
  cat("The number of outliers for", hotel_type, ":", length(hotel_df$adr[outliers]), "\n")
  # Outliers are stored in the data frame
  outliers_df <- rbind(outliers_df, hotel_df[outliers, ])
}

# Remove adr's outliers
df <- df[!(rownames(df) %in% rownames(outliers_df)), ] 
```


### data normalization

When we do data normalization, we only handle numerical data so we should firstly select the columns that need to be processed. That is to say, we need to delete data of Boolean type and char type.

It is worth noting that despite the company, agent, array_date_year, annual_date_ week_number, array_date_day_of_month is considered numerical data, but in reality they represent data of categorical or ordinal, so we do not consider normalizing this type of data.

```{r}
# Filter normalized data
# Filter data of numeric type
numeric_variables <- df[, sapply(df, is.numeric)]

# Filter binary data
non_binary_variables <- numeric_variables[, sapply(numeric_variables, 
                                                   function(col) length(unique(col)) > 2)]

# Filter data that have no practical significance
insignificance_values <- c("arrival_date_year", "arrival_date_week_number", 
                           "arrival_date_day_of_month","agent","company")
normalized_variables <- non_binary_variables[!(names(non_binary_variables) %in%
                                                 insignificance_values)]
```

For data with a small range of values, we believe that they exist within a reasonable range that satisfies the model's scale range for input data, so we do not intend to standardize them. Based on our analysis of the dataset, we plan to set this value threshold at 10.

```{r}
# Filter data that is less than 10 
max_values <- sapply(normalized_variables, max)
normalized_variables <- normalized_variables[, max_values > 10]
```

Having done the column filter process, we could check the the columns that are needed to do normalization. 

```{r}
# check data that should do normalization
colnames(normalized_variables)
```

We found that there are two commonly used data normalization methods, namely z-score and max min methods. We believe that each normalization method has some rationality, but here we choose to use the max-min normalization method for the following specific reasons.

1) Due to the handling of outliers in the previous steps, theoretically, there should be no obvious outliers or deviations in the data.

2) In addition, we did not find any dataset that satisfies a significant normal distribution or approximates a normal distribution.

So, we believe that using the max-min method here is more appropriate.

```{r}
# min-max
min_max <- function(x) {
  (x - min(x)) / (max(x) - min(x))
}

# min-max normalization
normalized_df <- as.data.frame(lapply(normalized_variables, min_max))
# Filter common column names and merge data
common_variables <- intersect(colnames(df), colnames(normalized_df))
df[, common_variables] <- normalized_df[, common_variables]
```


### Encode categorical values

Firstly, we select the attributes that need to be encoded, which are char type data, years and those that represent for IDs. There are a total of 12 attributes that need to be processed. We classify and encode these attributes based on the size and order of different variable categories.

```{r}
# Gain variables of char type
character_variables <- which(sapply(df, is.character))

# Convert the char type to the factor type
for (variable in character_variables) {
  df[, variable] <- as.factor(df[, variable])
}

# The number of variables of char type
unique_counts <- data.frame(Column = character(), Count = numeric())
for (variable in character_variables) {
  variable_name <- names(df)[variable] 
  unique_count <- length(unique(df[, variable]))
  unique_counts <- rbind(unique_counts, data.frame(Column = variable_name, 
                                                   Count = unique_count))
}

print(unique_counts)
```

For attributes of char types with a certain order, we choose to encode them using ordinal encoding, which is array_date_month.

```{r}
# Ordinal Encoding
factor_month <- factor(df$arrival_date_month, order = TRUE,
                       levels = c("January", "February", "March", 
                                  "April", "May", "June","July", 
                                  "August", "September", "October", 
                                  "November", "December"))

encode_month <- as.numeric(factor_month)
df$arrival_date_month <- encode_month
```

For attributes without ordinal order, we classify them into label encoding and one hot encoding. First of all, country, reserved_room_type and assigned_room_type, each attribute has many categories, for example, country has 116 categories, reserved_room_type has 8 categories, etc. If they are subjected to one hot encoding, the dimensionality of the data may be high, which may have a certain impact on the prediction results, which also have high computation cost. Therefore, for this type of data, we choose to use label encoding to avoid generating high-dimensional data.

```{r}
# Label Encoding
variables <- c("country","reserved_room_type", "assigned_room_type","agent","company")
for (encode_variable in variables) {
  df[[encode_variable]] <- as.integer(df[[encode_variable]])
}
```

For the remaining attributes, as their categories are not so many and there is no ordinal order, we have decided to directly adopt one hot encoding.

```{r}
# One-Hot Encoding
# Gain the variables' name of type factor
library(mltools)
library(data.table)
factor_variables <- names(df[sapply(df, is.factor)])
print(factor_variables)
for (encode_variable in factor_variables) {
  df <- one_hot(as.data.table(df), cols = encode_variable)
}
```

Since the company and agent here are IDs not numerical data, we believe that label encoding should also be performed.

```{r}
df$agent <- as.factor(df$agent)
df$company <- as.factor(df$company)
```

Finally, years should also be considered for encoding. We map different years onto a numerical label, and the final result is as follows.

```{r}
# encode year by mapping
df$arrival_date_year <- as.integer(factor(df$arrival_date_year, 
                                          levels = unique(df$arrival_date_year)))
```


### Store the preprocessed dataset

Finally, we store the processed data into a new_dataset.csv file, which contains 14374 rows and 48 columns.

```{r}
write.csv(df, file = "new_dataset.csv", row.names = FALSE)
```

# References
[1] N. Antonio, A. Almeida, L. Nunes, Hotel booking demand datasets. Data in Brief. 2019;22(41-49):41-49. doi:10.1016/j.dib.2018.11.126